---
title: "Week7"
author: "Brandon Cunningham"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(knitr)
library(jsonlite)
library(XML)
library(rvest)
library(httr)
```

## Introduction

In this assignment I will be creating 3 different data structures of some of my favorite books, Money ball, Red Notice, and Freakonomics.

First off I am using jsonlite to grab and read the json data from github.

```{r}
url_json <- 'https://raw.githubusercontent.com/btc2628/DATA607/main/Week%207%20Assignment/books.json'
jsonData <- jsonlite::fromJSON(url_json)
df_json <- as.data.frame(jsonData)
kable(df_json)
```

For XML I am first using httr to grab the data then XML to parse and transform it to a dataframe.

```{r}
xml_raw <- GET("https://raw.githubusercontent.com/btc2628/DATA607/main/Week%207%20Assignment/books.xml")
XMLData <- XML::xmlParse(xml_raw)
df_XML <- XML::xmlToDataFrame(XMLData)
kable(df_XML)
```
 Lastly I am using rvest to read in the html data and transform it into a dataframe.

```{r}
url_html<- 'https://raw.githubusercontent.com/btc2628/DATA607/main/Week%207%20Assignment/books.html'
htmlData <- rvest::read_html(url_html)
df_html <- rvest::html_table(htmlData, fill = TRUE)[[1]]
kable(df_html)
```

## Conlusion
This project has illustrated the different ways in which data can be formatted and strategies to transform that data to be consistent with dataframe structure for easy use.
