---
title: "Assignment 3"
author: "Brandon Cunningham"
date: "2024-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
library(readr)
library(dplyr)
library(tidyverse)
library(knitr)
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors <- read_csv(url)
```

#1 Find all majors that contain the word "Data" or "Statistics"

```{r}
data_or_stat <- majors[grepl("DATA|STATISTICS", majors$Major, ignore.case = TRUE), ]
data_or_stat$Major
```


#2 Transform The data

```{r, echo=FALSE}
text <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

text_no_ws = gsub('[^a-z"]', '', text)
text_comma = gsub('""', '", "', text_no_ws)
text_formatted <- paste0('c(', text_comma, ')')
cat(text_formatted)
```

#3 Explain the regex   
  
(.)\1\1: This regex would match a string of 3 of the same character. For instance it would match 'zzz' or '...'.  
"(.)(.)\\2\\1": This regex would match a 4 character string which are a palindrome. For instance it would match on 'xyyx' or '7;;7'.  
(..)\1: This regex would match any 4 characters where the first 2 characters and second 2 characters match. For instance it would match 'abab' or '2.2.'.  
"(.).\\1.\\1": This regex would match a 5 character string where the 1st, 3rd, and 5th character are the same character, and the 2nd and 4th can be any character. For instance it would match 'acaba' or '5:5m5'.  
"(.)(.)(.).*\\3\\2\\1": Would match a string of any length 6 characters or greater where the first 3 characters are repeated backwards as the last 3 characters. For instance it would match on 'abcn71';.,cba' or '123abc321'.  

#4 Construct regular expressions to match words that:  
  
Start and end with the same character: (.).*\\1  
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.): .\*(..).\*\\1.\*  
Contain one letter repeated in at least three places (e.g. "eleven" contains three e"s.): ".\*(.).\*\\1.\*\\1.\*

